---
title: "Data Science Fundamentals: Exercise"
author: "Samuel Chan"
date: "3/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
insurance <- read.csv("data_input/FL_insurance_sample.csv")
str(insurance)
```

The sample insurance file contains 36,634 records in Florida for 2012 from a sample company that implemented an agressive growth plan in 2012. There are total insured value (TIV) columns containing TIV from 2011 and 2012, so this dataset is great for testing out the comparison feature. This file has address information that you can choose to geocode, or you can use the existing latitude/longitude in the file.


```{r}
library(maps)
fl_fips <- county.fips[grep(x=county.fips$polyname, pattern="florida"), ]
fl_fips$polyname <- gsub("[[:alnum:]]+[,]", "", fl_fips$polyname)
head(fl_fips)
```


```{r}
counties_name <- tolower(gsub(pattern=" COUNTY", replacement = "", x = insurance$county, fixed = TRUE))
counties_name <- gsub(pattern="\\s+(?=[[:alnum:]])", " ", counties_name, perl = TRUE)
counties_name <- data.frame("counties_name" = counties_name)
counties_name$fips <- fl_fips[match(counties_name$counties_name, fl_fips$polyname),"fips"] 
head(counties_name)
```


```{r}
library(dplyr)
dat <- data.frame(
  "fips" = counties_name$fips,
  "y2012" = insurance$tiv_2012,
  "y2011" = insurance$tiv_2011
) %>% 
  group_by(fips) %>% 
  summarize_all(mean) 

head(dat)
```


```{r}
library(usmap)
library(ggplot2)
plot_usmap(data=dat, regions = "state", include="FL", values="y2011") + 
  scale_fill_gradient2(low='white', high='cadetblue', name="Total insured value") +
  labs(title="Total Insured Value by counties in Florida")
```

## Day 2: Catastrophe insurance
- Learn about catastrophe insurance
- Learning about subsetting
- Is there a more powerful way to generate cross-tables than table
- Learn about aggregating / summarizing data
- Introduction to machine learning

Total insured value is a term used to explain the total amount of insurance available in a single loss on a commercial property policy.  Total insured value is typically calculated by adding the property value, business interruption value and the value of any/all other property at the location. 

In simpler words: TIV describes the value of the building or the policy limit, whichever is smaller. 

A "natural disaster" insurance policy:
- Earthquake
- Hurricane
- Flood
- Fire

Q: Why not roll it into regular property insurance?
A: In the event of a hurricane hit or a massive flooding, all policyholders in a neighborhood will make a claim against the property insurance company - in fact it's nearly impossible to even determine the potential loss in such an event ahead of time. Unlike property damages covered in normal property insurance, catastrophic events are rarely isolated. 

Property insurance companies cannot afford the risk exposure as these could bankrupt them. They could increase the premium but they then wouldn't be competitive compared to other property insurance companies. Case in point: Florida, for instance, requires insurance companies to pay into a pool of reserve cash that is dedicated to protecting the policyholders of carriers unable to pay claims due to insolvency.

Consider Hurricane Andrew (1992) which resulted in USD 15 billion (adjusted for inflation, US$25 billion) in insurance claims, leading to the insolvency of 12 companies including:
- Regency Insurance Company  
- Great Republic Insurance Company
- Ocean Casualty Insurance Company

Read more: https://www.iii.org/article/spotlight-on-catastrophes-insurance-issues

Subset the first 10 rows of data, and columns 4 to 9:
```{r}
insurance[1:10,4:9]
```

```{r}
rowMax <- function(data) apply(data, 1, max, na.rm = TRUE)
rowMax(insurance[1:10,4:7])
```

```{r}
insurance[insurance$tiv_2011 > rowMax(insurance[,4:7]), 4:9]
```


## In-classroom Exercise
1. Are there any data cleansing required? 
  - Also: check for redundant features. Are both `line` and `construction` required for example?

2. Plot the relationship between TIV 2011 and TIV 2012? Do you think the insured value from the previous year could be a good indicator of insured value for the following year?

3. Use any exporatory techniques (for example: table, aggregate, xtabs, plot, ggplot) for the following questions?
  - 3a. Which county has the largest total insured value in 2012?
  - 3b. Which county has the greatest increase in total insured value from 2011 to 2012?
  - 3c. What features are highly correlated with our target (`tiv_2012`)? Knowing how to perform feature selection and extraction is  **critical part** of any machine learning projects 
  
